---
title: "hyperparameter-tuning"
format: 
  html:
    self-contained: true
editor: visual
author: "Claire Mittel"
---

# Data import/tidy/transform

```{r}
library(tidyverse)
library(powerjoin)
library(glue)
library(vip)
library(baguette)
library(tidymodels)
library(ggplot2)
library(ggthemes)
library(ranger)
library(xgboost)
library(dplyr)
library(skimr)
library(visdat)
library(rsample)
library(parsnip)
library(workflowsets)
library(workflows)
library(tune)
library(patchwork)
```

```{r}
root  <- 'https://gdex.ucar.edu/dataset/camels/file'
```

```{r}
download.file('https://gdex.ucar.edu/dataset/camels/file/camels_attributes_v2.0.pdf', 
              'data/camels_attributes_v2.0.pdf')
```

```{r}
types <- c("clim", "geol", "soil", "topo", "vege", "hydro")
```

```{r}
# Where the files live online ...
remote_files  <- glue('{root}/camels_{types}.txt')
# where we want to download the data ...
local_files   <- glue('data/camels_{types}.txt')
```

```{r}
walk2(remote_files, local_files, download.file, quiet = TRUE)
```

```{r}
camels <- map(local_files, read_delim, show_col_types = FALSE) 
```

```{r}
camels <- power_full_join(camels ,by = 'gauge_id')
```

```{r}
glimpse(camels)
skim(camels)
```

```{r}
vis_dat(camels)
vis_miss(camels)
```

```{r}
clean_data <- camels %>%
  janitor::clean_names() %>% 
  drop_na() %>%              
  filter(!is.na(gauge_lat) & !is.na(gauge_lon)) 
```

```{r}
ggplot(data = clean_data, aes(x = gauge_lon, y = gauge_lat)) +
  borders("state", colour = "gray50") +
  geom_point(aes(color = q_mean)) +
  scale_color_gradient(low = "pink", high = "dodgerblue") +
  ggthemes::theme_map()
```

# Data splitting

```{r}
set.seed(123)
```

```{r}
data_split <- initial_split(clean_data, prop = 0.80)  
train_data <- training(data_split)
test_data <- testing(data_split)
```

# Feature engineering

```{r}
q_recipe <- recipe(q_mean ~ ., data = train_data) %>%
  step_rm(gauge_lat, gauge_lon) %>%
  step_dummy(all_nominal_predictors(), -all_outcomes()) %>%
  step_impute_median(all_numeric_predictors()) %>%
  step_normalize(all_numeric_predictors())
```

```{r}
prepped_recipe <- prep(q_recipe)
baked_train <- bake(prepped_recipe, new_data = NULL)
```

# Resampling and model testing

```{r}
set.seed(123)  
folds <- vfold_cv(train_data, v = 10)
folds
```

```{r}
linear_spec <- linear_reg() %>%
  set_engine("lm") %>%
  set_mode("regression")
```

```{r}
rf_spec <- rand_forest(mtry = 3, trees = 500, min_n = 5) %>%
  set_engine("ranger") %>%
  set_mode("regression")
```

```{r}
boost_spec <- boost_tree(
  trees = 1000,
  tree_depth = 6,
  learn_rate = 0.01,
  loss_reduction = 0.01
) %>%
  set_engine("xgboost") %>%
  set_mode("regression")
```

```{r}
model_list <- list(
  linear_reg = linear_spec,
  random_forest = rf_spec,
  boosted_tree = boost_spec)

workflow_set <- workflow_set(
  preproc = list(q_recipe), 
  models = model_list)
```

```{r}
workflow_results <- workflow_map(
  workflow_set,
  resamples = folds,
  metrics = metric_set(rmse, rsq),  # can include mae if you want
  control = control_resamples(save_pred = TRUE)
)
```

```{r}
autoplot(workflow_results)
```

```{r}
collect_metrics(workflow_results)
```

I think that the boost tree model shows the best results for this data. I think this because it has the lowest RMSE and highest R sq value. The boost tree model uses the xgboost engine and is a regression model. I think that this model is working best because it tests nonlinear relationships.

# Model tuning

```{r}
tuned_boost_spec <- boost_tree(
  trees = 1000, 
  tree_depth = tune(),      
  learn_rate = tune(),      
  loss_reduction = tune()   
) %>%
  set_engine("xgboost") %>%
  set_mode("regression")
```

```{r}
tuned_boost_workflow <- workflow() %>%
  add_model(tuned_boost_spec) %>%
  add_recipe(q_recipe)
```

```{r}
dials <- extract_parameter_set_dials(tuned_boost_workflow)
dials$object
```

```{r}
my.grid <- grid_space_filling(
  x = dials,
  size = 25
)
```

```{r}
my.grid
```

```{r}
model_params <- tune_grid(
  tuned_boost_workflow,      
  resamples = folds,          
  grid = my.grid,             
  metrics = metric_set(rmse, rsq, mae),
  control = control_grid(save_pred = TRUE)
)

autoplot(model_params)

```

What I see is that the results for each of the models is pretty scattered but the most scattered is the middle one.

```{r}
collect_metrics(model_params) %>%
  arrange(mean)
```

```{r}
show_best(model_params, metric = "mae", n = 5)
```

I see that the best hyperparameter is tree depth from this table.

```{r}
hp_best <- select_best(model_params, metric = "mae")
```

```{r}
final_wf <- finalize_workflow(
  tuned_boost_workflow, 
  hp_best)
```

# Final model verification

```{r}
final_fit <- last_fit(
  final_wf,
  split = data_split
)
```

```{r}
collect_metrics(final_fit)
```

```{r}
final_preds <- collect_predictions(final_fit)
```

```{r}
ggplot(final_preds, aes(x = .pred, y = q_mean)) +
  geom_point(aes(color = q_mean), alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE, color = "red", linetype = "dashed") +
  geom_abline(slope = 1, intercept = 0, color = "navy", linetype = "solid") +
  scale_color_viridis_c(option = "plasma") +
  labs(
    title = "Predicted vs. Actual q_mean on Test Set",
    x = "Predicted q_mean",
    y = "Actual q_mean",
    color = "True q_mean"
  ) +
  theme_minimal()
```

# Building a map

```{r}
final_model <- fit(final_wf, data = clean_data)
```

```{r}
pred_data <- augment(final_model, new_data = clean_data)
```

```{r}
pred_data <- pred_data %>%
  mutate(residual_sq = (q_mean - .pred)^2)
```

```{r}
map_preds <- ggplot(data = pred_data, aes(x = gauge_lon, y = gauge_lat)) +
  borders("state", colour = "gray50") +
  geom_point(aes(color = .pred)) +
  scale_color_gradient(low = "pink", high = "dodgerblue") +
  ggthemes::theme_map() +
  labs(title = "Predicted q_mean Across CONUS", color = "Predicted q_mean")
```

```{r}
map_resid <- ggplot(data = pred_data, aes(x = gauge_lon, y = gauge_lat)) +
  borders("state", colour = "gray50") +
  geom_point(aes(color = residual_sq)) +
  scale_color_gradient(low = "pink", high = "dodgerblue") +
  ggthemes::theme_map() +
  labs(title = "Squared Residuals Across CONUS", color = "Residuals²")
```

```{r}
map_preds + map_resid + plot_layout(ncol = 2)
```
